---
title: "Mini-Project #01: What We Watched"
author: "Konstantinos Gkanas" 
class: STA 9750 - Software Tools for Data Analysis
format:
  html:
    theme: darkly
    css: styles.css
    toc: true
---

```{css, echo=FALSE}
/* Netflix Red Accent Styling */
body {
    background-color: #141414 !important;
    color: #ffffff !important;
}

a {
    color: #E50914 !important;
}

h1, h2, h3 {
    color: #ffffff !important;
    border-bottom: 2px solid #E50914;
    padding-bottom: 5px;
}

.callout.callout-note {
    background-color: #1f1f1f !important;
    border-left: 5px solid #E50914 !important;
}
---

# Data Acquisition & Cleaning
```{r}
#| message: false
#| warning: false
#| echo: false

library(tidyverse)
library(rvest)
library(readxl)
library(fs)
library(gt)

load_netflix_data <- function(){
  data_dir <- path("data", "mp01")
  
  if(!dir_exists(data_dir)){
    dir_create(data_dir, showWarnings=FALSE, recurse=TRUE)
  }
  
  NETFLIX_URLS <- list(
    `2023H1` = "https://about.netflix.com/news/what-we-watched-a-netflix-engagement-report",
    `2023H2` = "https://about.netflix.com/news/what-we-watched-the-second-half-of-2023",
    `2024H1` = "https://about.netflix.com/news/what-we-watched-the-first-half-of-2024",
    `2024H2` = "https://about.netflix.com/news/what-we-watched-the-second-half-of-2024",
    `2025H1` = "https://about.netflix.com/news/what-we-watched-the-first-half-of-2025",
    `2025H2` = "https://about.netflix.com/news/what-we-watched-the-second-half-of-2025",
    `2026H1` = "https://about.netflix.com/news/what-we-watched-the-second-half-of-2026"
  )
  
  download_if_needed <- function(url, local){
    if(!file_exists(local)){
      download.file(url, local, quiet = TRUE, mode = "wb")
    }
    return(local)
  }
  
  NETFLIX_URLS |> 
    map(read_html) |> 
    map(html_element, xpath = "//a[contains(@href, '.xlsx')]") |> 
    map_chr(html_attr, "href") |> 
    enframe("period", "url") |>
    mutate(path = path(data_dir, paste0(period, ".xlsx")), 
           path = map2(url, path, download_if_needed), 
           data = map(path, read_excel, skip=3, na="*"), 
           data = map(data, \(x) x |> mutate(`Release Date` = as.Date(`Release Date`)))) |>
    unnest(data) |> 
    select(-url, -path) |> 
    mutate(year=str_sub(period, 1, 4), 
           half=str_sub(period, start=6), 
           period_start = make_date(year, if_else(half==1, 1, 7), 1), 
           period_end = make_date(year, if_else(half==1, 6, 12), if_else(half==1, 30, 31))) |> 
    filter(!is.na(`Hours Viewed`)) |> 
    mutate(live_event = !is.na(Runtime) & str_detect(Runtime, "\\d+:\\d+\\*"), 
           clean_time = str_remove(Runtime, fixed("*")),
           Runtime = 60 * hour(hm(clean_time)) + minute(hm(clean_time))) |> 
    select(-clean_time) |> 
    mutate(Runtime = coalesce(Runtime, mean(Runtime, na.rm=TRUE)), 
           .by=Title, 
           Title = stringi::stri_trans_general(Title, "Latin-ASCII"),
           Title = str_remove(Title, " // .*"), 
           Title = str_remove(Title, "[^[:ascii:]].*"), 
           Title = str_trim(Title)) |> 
    filter_out(Title == "") |> 
    rename(
      title = Title,
      runtime = Runtime,
      views = Views,
      hours_viewed = `Hours Viewed`,
      release_date = `Release Date`,
      available_globally = `Available Globally?`
    ) |>
    mutate(
      year = as.integer(year),
      available_globally = (available_globally == "Yes")
    ) |>
    select(year, period, period_start, period_end, title, runtime, views, 
           hours_viewed, release_date, available_globally, live_event)
}

NETFLIX_DATA <- load_netflix_data()
```

# Further Analysis & Tables 
```{r}
#| label: tbl-consistent-staples
#| tbl-cap: "Top Consistent 'Staple' Titles Across Multiple Periods"
#| echo: false
#| message: false
#| warning: false

NETFLIX_DATA |>
  group_by(title) |>
  summarize(
    appearances = n(),
    avg_hours = mean(hours_viewed, na.rm = TRUE),
    min_hours = min(hours_viewed, na.rm = TRUE),
    sd_hours = sd(hours_viewed, na.rm = TRUE)
  ) |>
  filter(appearances >= 5) |> 
  arrange(desc(min_hours)) |>
  slice_head(n = 5) |>
  gt() |>
  fmt_number(columns = c(avg_hours, min_hours, sd_hours), decimals = 0) |>
  cols_label(
    title = "Title",
    appearances = "Periods Appeared",
    avg_hours = "Average Hours",
    min_hours = "Minimum Hours",
    sd_hours = "Std. Deviation"
  )
```

# Shows vs. Movies 
```{r}
#| label: tbl-shows-vs-movies
#| tbl-cap: "Fraction of Watch Time: Shows vs. Movies"
#| echo: false
#| message: false
#| warning: false

NETFLIX_DATA |>
  mutate(
    content_type = if_else(runtime > 240 | str_detect(title, "(?i)Season"), "TV Show", "Movie")
  ) |>
  group_by(period, content_type) |>
  summarize(total_hours = sum(hours_viewed, na.rm = TRUE), .groups = "drop") |>
  group_by(period) |>
  mutate(fraction = total_hours / sum(total_hours)) |>
  select(period, content_type, fraction) |>
  pivot_wider(names_from = content_type, values_from = fraction) |>
  gt() |>
  fmt_percent(columns = c(`TV Show`, Movie), decimals = 1) |>
  cols_label(period = "Period")
```

# Live Events Vs. Pre-Recorded Content
```{r}
#| label: tbl-live-vs-taped
#| tbl-cap: "Performance Comparison: Live Events vs. Pre-Taped Content"
#| echo: false
#| message: false
#| warning: false

NETFLIX_DATA |>
  mutate(content_format = if_else(live_event, "Live Event", "Pre-Taped")) |>
  group_by(content_format) |>
  summarize(
    total_titles = n(),
    avg_hours = mean(hours_viewed, na.rm = TRUE),
    max_hours = max(hours_viewed, na.rm = TRUE)
  ) |>
  gt() |>
  fmt_number(columns = c(avg_hours, max_hours), decimals = 0) |>
  cols_label(
    content_format = "Format",
    total_titles = "Total Offerings",
    avg_hours = "Average Viewership (Hrs)",
    max_hours = "Top Tier Viewership (Hrs)"
  )
```